*Supplementary Material 1*

### 2022-02-26 ms/manuscript.Rmd
WIP: add bs values

tar_load(
  c(
    accs_exclude,
    fern_sanger_seqs_raw,
    fossil_calibration_tips,
    monophy_by_clade,
    plastid_calibration_dates,
    plastid_tree_dated,
    plastome_alignment,
    plastome_alignment_tbl,
    plastome_metadata_raw_all,
    plastome_metadata_renamed,
    plastome_metadata_renamed,
    plastome_seqs_combined_filtered,
    plastome_tree,
    ppgi_taxonomy,
    pteridocat,
    raw_fasta_all,
    sanger_accessions_selection,
    sanger_alignment,
    sanger_alignment_tbl,
    sanger_seqs_combined_filtered,
    sanger_seqs_rogues_removed,
    treepl_cv_results,
    ts_fossil_calibration_tips
  ),
  store = ftol_cache
)

@@

@@

@@

We estimate the crown age of ferns to be `r crown_ages$ferns` ma.
Ages of other major crown groups are leptosporangiates (`r crown_ages$leptos` ma), Polypodiales (`r crown_ages$polypodiales` ma), eupolypods I (`r crown_ages$eupoly_i` ma), and eupolypods II (`r crown_ages$eupoly_ii` ma).
Except for crown ferns, these ages are all significantly older than those reported in the most recent global fern phylogeny [@Testo2016a].
Estimated stem ages of fern families were also older than previous studies (`r figure("divtimes")`).
We did not compare crown ages of families across studies because differences in crown ages in smaller clades are likely affected by species sampling.

>>>>

tar_load(
  c(
    accs_exclude,
    fern_sanger_seqs_raw,
    fossil_calibration_tips,
    monophy_by_clade,
    plastid_calibration_dates,
    plastid_tree_dated,
    plastome_alignment,
    plastome_alignment_tbl,
    plastome_metadata_raw_all,
    plastome_metadata_renamed,
    plastome_metadata_renamed,
    plastome_seqs_combined_filtered,
    plastome_tree,
    ppgi_taxonomy,
    pteridocat,
    raw_fasta_all,
    sanger_accessions_selection,
    sanger_alignment,
    sanger_alignment_tbl,
    sanger_con_tree_rooted,
    sanger_seqs_combined_filtered,
    sanger_seqs_rogues_removed,
    treepl_cv_results,
    ts_fossil_calibration_tips
  ),
  store = ftol_cache
)

@@

# Mean bootstrap for consensus tree
tibble(
  bs = sanger_con_tree_rooted$node.label)

@@

Bootstrap support was generally across the tree (mean BS) and particularly high at deeper nodes.
Relationships within some more recently diverged clades were less well-supported, including *Cyathea*, *Amauropelta*, and parts of *Dryopteris* and *Elaphoglossum* (`r s_figure("bs")`).

@@

We estimate the crown age of ferns to be `r crown_ages$ferns` ma.
Ages of other major crown groups are leptosporangiates (`r crown_ages$leptos` ma), Polypodiales (`r crown_ages$polypodiales` ma), eupolypods I (`r crown_ages$eupoly_i` ma), and eupolypods II (`r crown_ages$eupoly_ii` ma).
Our estimates for leptosporangiates and Polypodiales are both `r ca` 50 ma older than the most recent global fern phylogeny [@Testo2016a], while other ages are similar.
Estimated stem ages of fern families were mostly older than previous studies (`r figure("divtimes")`).
We did not compare crown ages of families across studies because differences in crown ages in smaller clades are likely affected by species sampling.

### 2022-02-26 ms/si.Rmd
WIP: add bs values

tar_load(
  c(
    sanger_alignment_tbl,
    sanger_sampling,
    sanger_accessions_selection,
    plastome_tree,
    plastid_tree_dated,
    fossil_calibration_tips,
    ppgi_taxonomy,
    monophy_by_clade,
    acc_table_long,
    plastome_seqs_combined_filtered,
    sanger_tree_rooted,
    fossil_calibrations_for_treepl,
    sanger_tree_dated,
    fossil_ferns_all),
  store = ftol_cache
)

@@

>>>>

tar_load(
  c(
    sanger_alignment_tbl,
    sanger_sampling,
    sanger_accessions_selection,
    plastome_tree,
    plastid_tree_dated,
    fossil_calibration_tips,
    ppgi_taxonomy,
    monophy_by_clade,
    acc_table_long,
    plastome_seqs_combined_filtered,
    sanger_tree_rooted,
    fossil_calibrations_for_treepl,
    sanger_tree_dated,
    sanger_con_tree_rooted,
    fossil_ferns_all),
  store = ftol_cache
)

@@

```{r bs-plot}
# Parse bootstrap values from IQTREE
phy <- sanger_con_tree_rooted

bs_values <-
tibble(
  label_orig = phy$node.label
) %>%
  mutate(label_new = case_when(
    label_orig == "Root" ~ NaN,
    TRUE ~ parse_number(label_orig) * 0.01
  )) %>%
  pull(label_new)

# Make tibble for printing BS values
bs_tibble <- tibble(
  # hard-code node ID: internal nodes start after tip nodes,
  # and phy$node.label is in the same order as internal nodes
  node = 1:Nnode(phy) + Ntip(phy), 
  # Don't print BS = 100%
  bootstrap = bs_values)

# Make plot
# optional: round version
bs_plot_round <- 
  ggtree(
    phy,
    layout = "circular",
    branch.length = "none", 
    size = 0.1, color = "grey30") %<+% bs_tibble +
  # Add node points
  geom_nodepoint(
    aes(color = bootstrap), size = 0.2) +
  scale_color_viridis_c(
    "BS",
    option = "B", na.value = "transparent",
    breaks = 0.25*0:4, labels = percent(0.25*0:4))

# Make plot
bs_plot <- 
  ggtree(
    phy,
    size = 0.1, color = "grey30") %<+% bs_tibble +
  # Add node points
  geom_nodepoint(
    aes(color = bootstrap), size = 0.1) +
  scale_color_viridis_c(
    "BS",
    option = "B", na.value = "transparent",
    breaks = 0.25*0:4, labels = percent(0.25*0:4)) +
  # Add tip labels
  geom_tiplab(aes(label = label), size = 0.5, color = "grey30") +
  theme(legend.position = "bottom")
```

```{r bs-save, eval = params$doc_type == "doc"}
ggsave(
  plot = bs_plot,
  file = (result_file(s_figure("bs"), "pdf")),
  height = 60, width = 8, units = "in",
  limitsize = FALSE)
```

**`r s_figure("bs")`**.
Consensus fern phylogeny with node labels bootstrap support (%) (`r result_file(s_figure("bs"), "pdf") %>% fs::path_file()`).
